Stateless Model Checking with Data-Race Preemption Points

### 2016-07-19 oopsla-16/overview.tex
fix some camready todo layout issues now that we are swimming in free space
Philosophy of bugs.

\begin{figure}[t]
	\small
\begin{tabular}{c}
	\revision{
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~// A1} & \\
	2 & \texttt{mutex\_lock(\&mx);} & \\
	3 & \texttt{mutex\_unlock(\&mx);} & \\
	4 & & \texttt{mutex\_lock(\&mx);} \\
	5 & & \texttt{mutex\_unlock(\&mx);} \\
	6 & & \texttt{\hilight{brickred}{x++;}~// A2} \\
\end{tabular}
}
\\
\revision{\normalsize (a) True potential data race.}
% TODO CAMREADY: put another line break here
\\
\revision{
\begin{tabular}{rll}
	%& \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
	%& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~// B1} & \\
	2 & \texttt{mutex\_lock(\&mx);} & \\
	3 & \texttt{y = true;} & \\
	4 & \texttt{mutex\_unlock(\&mx);} & \\
	5 & & \texttt{mutex\_lock(\&mx);} \\
	6 & & \texttt{bool tmp = y;} \\
	7 & & \texttt{mutex\_unlock(\&mx);} \\
	8 & & \texttt{if (tmp) \hilight{brickred}{x++;}~// B2} \\
	%8 & & \texttt{if (tmp)} \\
	%9 & & \texttt{~~~~\hilight{brickred}{x++;}~// B2} \\
\end{tabular}
}
\\
\revision{\normalsize (b) No data race in any interleaving.}
\end{tabular}
\caption{\revision{Data-race analyses may be prone to either {\em false negatives} or {\em false positives}.
Applying HB to program (a) will miss the potential race possible between A1/A2 in an alternate interleaving,
while using Limited HB on (b) will produce a false alarm on B1/B2.}}
\label{fig:hb-example}
\end{figure}

@@

>>>>

\begin{figure}[t]
	\small
\begin{tabular}{c}
	\revision{
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~// A1} & \\
	2 & \texttt{mutex\_lock(\&mx);} & \\
	3 & \texttt{mutex\_unlock(\&mx);} & \\
	4 & & \texttt{mutex\_lock(\&mx);} \\
	5 & & \texttt{mutex\_unlock(\&mx);} \\
	6 & & \texttt{\hilight{brickred}{x++;}~// A2} \\
\end{tabular}
}
\\
\revision{\normalsize (a) True potential data race.}
\\
\\
\revision{
\begin{tabular}{rll}
	%& \multicolumn{2}{c}{\texttt{int x = 0; bool y = false; mutex\_t mx;}} \\
	%& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{brickred}{x++;}~// B1} & \\
	2 & \texttt{mutex\_lock(\&mx);} & \\
	3 & \texttt{y = true;} & \\
	4 & \texttt{mutex\_unlock(\&mx);} & \\
	5 & & \texttt{mutex\_lock(\&mx);} \\
	6 & & \texttt{bool tmp = y;} \\
	7 & & \texttt{mutex\_unlock(\&mx);} \\
	8 & & \texttt{if (tmp) \hilight{brickred}{x++;}~// B2} \\
	%8 & & \texttt{if (tmp)} \\
	%9 & & \texttt{~~~~\hilight{brickred}{x++;}~// B2} \\
\end{tabular}
}
\\
\revision{\normalsize (b) No data race in any interleaving.}
\end{tabular}
\caption{\revision{Data-race analyses may be prone to either {\em false negatives} or {\em false positives}.
Applying HB to program (a) will miss the potential race possible between A1/A2 in an alternate interleaving,
while using Limited HB on (b) will produce a false alarm on B1/B2.}}
\label{fig:hb-example}
\end{figure}

@@

\revisionx{
{\bf Philosophy of bugs.}
% two camps: all races are bugs; citations:
% and, races must be classified; citations: replay analysis, portend,
While there is a vast body of work on how to detect data races to begin with,
judging data races once found is a matter of philosophical debate unto itself.
Some recent tools classify races depending on how they impact program behaviour \cite{recordreplaydrs,portend}, overlooking {\em benign} races in search of more dangerous ones.
%In particular,
\cite{racerx} acknowledges that a program may have too many races for a user to worry about,
so bug reports must be prioritized by severity of effects. %how likely they may cause failures or security vulnerabilities.
%
However, other prior work argues that data races are always bugs \cite{miscompile-benign,data-races-are-evil},
largely due to the possibility of compiler or hardware reordering of racy accesses.
%
We take the former camp:
we consider a data race a bug only when it results in a visible failure state (e.g., crash or deadlock).
We bypass concerns of compiler reordering by checking programs at the executable level;
for a study of hardware reordering in the context of DPOR, we refer the reader to \cite{tsopso}.
}
